---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


``````{r}
load("./data_clean_time_result.RData")
```

```{r}
players = read.csv("./Players.csv")
```

```{r}
players = players %>% mutate(LastPlayedDate2 = ifelse(is.na(LastPlayedDate), "2022-3-6", LastPlayedDate))
```

```{r}
players = players %>% mutate(
time_in_MLB = as.numeric(difftime(as.Date(players$LastPlayedDate2), as.Date(players$MLBDebut), units = "days"))/365.25)
```

```{r}
players = players %>% mutate(
time_in_MLB2 = ifelse(is.na(time_in_MLB
), 0, time_in_MLB))
```

```{r}
players = players %>% mutate(
success = time_in_MLB2 > 5, TRUE, FALSE)
```


```{r}
l1 = lm(Career_WAR ~ success, data = players)
```


```{r}
anova(l1)
```
```{r}
t.test(Career_WAR ~ success, data = players, alternative = "less")
```

```{r}
data_1 = merge(data_clean_time_single, players, by = "RegLookup")
```


```{r}
boxplot(players$Career_WAR~players$success)
```


```{r}
table(data_1$is_miss_10_months, data_1$success)
```

