---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


``````{r}
library(dplyr)
library(ggplot2)
library(scales)
load("./data_clean_time_result.RData")
```

```{r}
players = read.csv("./Players.csv")
```

```{r}
players = players %>% mutate(LastPlayedDate2 = ifelse(is.na(LastPlayedDate), "2022-3-6", LastPlayedDate))
```

```{r}
players = players %>% mutate(
time_in_MLB = as.numeric(difftime(as.Date(players$LastPlayedDate2), as.Date(players$MLBDebut), units = "days"))/365.25)
```

```{r}
players = players %>% mutate(
time_in_MLB2 = ifelse(is.na(time_in_MLB
), 0, time_in_MLB))
```

```{r}
players = players %>% mutate(
success = time_in_MLB2 > 5, TRUE, FALSE)
```

```{r}
#removing designated hitters
players <- players[-which(players$PrimaryPos == "DH"),]
```


```{r}
load("data_clean_time_single.RData")
load("data_clean_time_single_3_months.RData")
#data_1 = merge(data_clean_time_single, players, by = "RegLookup")
data_1 = merge(data_clean_time_single_3_months, players, by = "RegLookup")
```

```{r}
#unique(data_1$PrevType)
data_1_4yr <- data_1 %>%
  filter(PrevType == "4Yr")

data_1_hs <- data_1 %>%
  filter(PrevType == "HS")

data_1_jc <- data_1 %>%
  filter(PrevType == "JC")
```


```{r}
boxplot_4yr <- ggplot(data_1_4yr, aes(x=success, y=Career_WAR, fill=success)) + 
  geom_boxplot(lwd=0.2) +
  ylim(0, 80) +
  scale_fill_manual(values=c("#d95543", "#39994f")) +
  labs(title = "Career WAR distribution for Players Drafted in 4 Year Colleges", x="Player was Successful?", y="Career WAR") +
  theme(text=element_text(family = "Times New Roman"), plot.title = element_text(face="bold"), legend.position = "none")
boxplot_4yr

```


```{r}
boxplot_hs <- ggplot(data_1_hs, aes(x=success, y=Career_WAR, fill=success)) + 
  geom_boxplot(lwd=0.2) +
  ylim(0, 80) +
  scale_fill_manual(values=c("#d95543", "#39994f")) +
  labs(title = "Career WAR distribution for Players Drafted in High School", x="Player was Successful?", y="Career WAR") +
  theme(text=element_text(family = "Times New Roman"), plot.title = element_text(face="bold"), legend.position = "none")
boxplot_hs
```


```{r}
boxplot_jc <- ggplot(data_1_jc, aes(x=success, y=Career_WAR, fill=success)) + 
  geom_boxplot(lwd=0.2) +
  ylim(0, 80) +
  scale_fill_manual(values=c("#d95543", "#39994f")) +
  labs(title = "Career WAR distribution for Players Drafted in Junior College", x="Player was Successful?", y="Career WAR") +
  theme(text=element_text(family = "Times New Roman"), plot.title = element_text(face="bold"), legend.position = "none")
boxplot_jc
```


```{r}
t.test(Career_WAR ~ success, data = data_1_4yr, alternative = "less")
t.test(Career_WAR ~ success, data = data_1_hs, alternative = "less")
t.test(Career_WAR ~ success, data = data_1_jc, alternative = "less")
```

```{r}
t.test(Career_WAR ~ is_miss_3_months_season, data = data_1_4yr, alternative = "greater")
t.test(Career_WAR ~ is_miss_3_months_season, data = data_1_hs, alternative = "greater")
t.test(Career_WAR ~ is_miss_3_months_season, data = data_1_jc, alternative = "greater")
```

```{r}
t.test(Career_WAR ~ is_miss_10_months, data = data_1_4yr, alternative = "greater")
t.test(Career_WAR ~ is_miss_10_months, data = data_1_hs, alternative = "greater")
t.test(Career_WAR ~ is_miss_10_months, data = data_1_jc, alternative = "greater")
```


```{r}
l_4yr = lm(Career_WAR ~ success, data = data_1_4yr)
anova(l_4yr)

l_hs = lm(Career_WAR ~ success, data = data_1_hs)
anova(l_hs)

l_jc = lm(Career_WAR ~ success, data = data_1_jc)
anova(l_jc)
```

```{r}
#updating positions
data_2 <- data_1

data_2[which(data_2$PrimaryPos == "1B" | data_2$PrimaryPos == "2B" | data_2$PrimaryPos == "3B" | data_2$PrimaryPos == "SS" | data_2$PrimaryPos == "IF"),]$PrimaryPos <- "IF"

data_2[which(data_2$PrimaryPos == "LF" | data_2$PrimaryPos == "OF" | data_2$PrimaryPos == "RF" | data_2$PrimaryPos == "CF"),]$PrimaryPos <- "OF"

data_2_4yr <- data_2 %>%
  filter(PrevType == "4Yr")

data_2_hs <- data_2 %>%
  filter(PrevType == "HS")

data_2_jc <- data_2 %>%
  filter(PrevType == "JC")

```


anova(Career WAR ~ is_miss_10_months + position)

```{r}

#l2 = lm(Career_WAR ~ is_miss_3_months_season + PrimaryPos, data = data_2)
#anova(l2)

l_4yr_2 = lm(Career_WAR ~ is_miss_3_months_season + PrimaryPos, data = data_2_4yr)
anova(l_4yr_2)
summary(l_4yr_2)


l_hs_2 = lm(Career_WAR ~ is_miss_3_months_season + PrimaryPos, data = data_2_hs)
anova(l_hs_2)
summary(l_hs_2)

l_jc_2 = lm(Career_WAR ~ is_miss_3_months_season + PrimaryPos, data = data_2_jc)
anova(l_jc_2)
summary(l_jc_2)
```

```{r}
l_4yr_3 = lm(Career_WAR ~ is_miss_10_months + PrimaryPos, data = data_2_4yr)
anova(l_4yr_3)
summary(l_4yr_3)


l_hs_3 = lm(Career_WAR ~ is_miss_10_months + PrimaryPos, data = data_2_hs)
anova(l_hs_3)
summary(l_hs_3)

l_jc_3 = lm(Career_WAR ~ is_miss_10_months + PrimaryPos, data = data_2_jc)
anova(l_jc_3)
summary(l_jc_3)
```


```{r}


data_1 <- data_1 %>%
  filter(PrevType %in% c("4Yr", "HS", "JC"))
boxplot_bonus <- ggplot(data_1, aes(x=PrevType, y=Bonus, fill=PrevType)) + 
  geom_boxplot(lwd=0.2) + 
  scale_fill_manual(values=c("#d95543", "#39994f", "#98bcdc")) +
  labs(title = "Bonus distribution for Draft Time Groups", x="Draft Time", y="Bonus (Millions)") +
  scale_x_discrete(labels = c("4 Year College", "High School", "Junior College")) +
  scale_y_continuous(labels = label_number(suffix = "M", scale = 1e-6)) +
  theme(text=element_text(family = "Times New Roman"), plot.title = element_text(face="bold"), legend.position = "none")

boxplot_bonus
```


```{r}
data_hs_4yr <- data_1 %>%
  filter(PrevType %in% c("4Yr", "HS"))

data_hs_jc <- data_1 %>%
  filter(PrevType %in% c("HS", "JC"))

t.test(Bonus ~ PrevType, data = data_hs_4yr, alternative = "less")
t.test(Bonus ~ PrevType, data = data_hs_jc, alternative = "greater")
```


