---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r}
load("./data_clean_time_result.RData")
```

```{r}
load("data_clean_time_single.RData")
```

```{r}
data_clean_time5 = list()
for (i in 1:714){
  data_i = data_clean_time_result[[i]]
  data_result = data_i %>% filter(as.numeric(time_difference) > 90)
  data_clean_time5 = append(data_clean_time5, list(data_result))
}
```

```{r}
names(data_clean_time5) = names(data_clean_time_result)
```

```{r}
data_clean_time6 = list()
for(i in 1:714){
  data_i = data_clean_time5[[i]]
  data_result = data_i %>% mutate(year_to_day = yday(Date), current_season_miss = yday(Date) - 90, prev_year_miss = as.numeric(time_difference) - year_to_day - 92)
  data_result = data_result %>% mutate(miss_total = prev_year_miss + current_season_miss)
  data_result = data_result %>% mutate(is_miss_3_months_season = miss_total > 90)
  data_clean_time6 = append(data_clean_time6, list(data_result))
}
```

```{r}
data_clean_time6[[714]]
```

```{r}
miss_3_months = c()
for(i in 1:714){
  if(nrow(data_clean_time_result[[i]]) == 0)
  {
    miss_3_months = append(miss_3_months, NA)
  }
  else{
    
    miss_3_months = append(miss_3_months, sum(data_clean_time6[[i]]$is_miss_3_months_season) > 0)
  }
}
```

```{r}
data_clean_time_single
```


```{r}
table(miss_3_months)
```
```{r}
data_clean_time_result[[41]]
```


```{r}
table(data_clean_time_single$is_miss_10_months)
```
```{r}
data_clean_time_single_3_months = data_clean_time_single %>% mutate(is_miss_3_months_season = miss_3_months)
```

```{r}
save(data_clean_time_single_3_months, file = "data_clean_time_single_3_months.RData")
```

